---
title: "Case_Study_1"
author: "Karl J. Jurek"
date: "December 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
# Loading the 'Beers.csv' file into the 'beers' dataframe
beers=read.csv("Beers.csv", header=T, stringsAsFactors=F)

# Loading the 'Breweries.csv' file into the 'breweries' dataframe
breweries=read.csv("Breweries.csv", header=T, stringsAsFactors=F)
```

## Question 1: How many breweries are present in each state?

```{r Question_1, echo=TRUE}
# Adding the 'plyr' library
library('plyr')

# Creating a count of all facilities in each state in the 'count_of_breweries_by_state' data frame
count_of_breweries_by_state=ddply(breweries, .(breweries$State), nrow)

# Assigning useful names to the columns / variables in the 'count_of_breweries_by_state' data frame
names(count_of_breweries_by_state)=c("State", "Count")

# Displaying the count of breweries by state
head(count_of_breweries_by_state)

# Displaying the count of breweries by Count
head(arrange(count_of_breweries_by_state, desc(Count)))

```

## Question 2A: Merge beer data with the breweries data. 

```{r Question_2A, echo=TRUE}
# Combining the 'beers' and 'breweries' dataframes into the 'bandb' dataframe
bandb=merge(beers, breweries, by=c("Brewery_id"), by.y=c("Brew_ID"), all=T)

# Removing extra space from the 'State' column in the 'bandb' dataframe
bandb$State=gsub(" ", "", bandb$State)

# Assigning useful name to the 'Name.x' and 'Name.y' columns
names(bandb)[2]="BeerName"
names(bandb)[8]="Brewery"
```


## Question 2B: Print the first 6 observations and the last six observations to check the merged file.

```{r Question_2B, echo=TRUE}
# Displaying the first six rows of the merged dataframe, 'bandb'
head(bandb)

# Displaying the last six rows of the merged dataframe, 'bandb'
tail(bandb)
```

## Question 3: Report the number of NA's in each column.

```{r Question_3, echo=TRUE}
colSums(is.na(bandb))
```

## Question 4A: Compute the median alcohol content (ABV) for each state.
```{r Question_4A, echo=TRUE}
# Adding 'dplr' library to project
library('dplyr')

# Getting the median ABV for each state
median_abv_by_state=bandb %>% group_by(bandb$State) %>% summarize(medianabv=median(ABV, na.rm=T))

# Assigning useful name to the 'median_abv_by_state' dataframe
names(median_abv_by_state)=c("State", "medianABV")

# Displaying partial results for the 'median_abv_by_state' dataframe by State
head(median_abv_by_state)

# Displaying partial results for the 'median_abv_by_state' dataframe by median ABV
head(arrange(median_abv_by_state, desc(medianABV)))
```

## Question 4B: Compute the median international bitterness unit (IBU) for each state.
```{r Question_4B, echo=TRUE}
# Getting the median IBU for each state
median_ibu_by_state=bandb %>% group_by(bandb$State) %>% summarize(medianibu=median(IBU, na.rm=T))

# Assigning useful names to the 'median_ibu_by_state' dataframe
names(median_ibu_by_state)=c("State", "medianIBU")

# Displaying partial results for the 'median_ibu_by_state' dataframe by State
head(median_ibu_by_state)

# Displaying partial results for the 'median_ibu_by_state' dataframe by median IBU
head(arrange(median_ibu_by_state, desc(medianIBU)))
```

## Question 4C: Plot a bar chart to compare alcohol content (ABV) for each state.
```{r Question_4C, fig.width=10,fig.height=11, echo=TRUE}
# Adjusted the size of the plot with [fig.width=10,fig.height=11,] in the code chunk for this section

# Adding the 'ggplot2' library to the project
library('ggplot2')

# Creating bar graph of the 'median_abv_by_state' dataframe
ggplot(median_abv_by_state, aes(reorder(State, medianABV), y=medianABV, fill=medianABV)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=medianABV), hjust=1.2, vjust=0.5, color="white", size=2.5) +
  ggtitle("Median ABV by State") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x="Alcohol by Volume (ABV)", y="States", fill="medianABV") + 
  coord_flip() + 
  guides(fill=FALSE)
```

## Question 4D: Plot a bar chart to compare international bitterness unit (IBU) for each state.
```{r Question_4D, fig.width=10,fig.height=11, echo=TRUE}
# Adjusted the size of the plot with [fig.width=10,fig.height=11,] in the code chunk for this section

# Removing rows with 'NA' values
no_na_median_ibu=data.frame(median_ibu_by_state[!is.na(median_ibu_by_state$medianIBU), ])

# Creating bar graph of the 'no_na_median_ibu' dataframe
ggplot(no_na_median_ibu, aes(reorder(State, medianIBU), y=medianIBU, fill=medianIBU)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=medianIBU), hjust=1.2, vjust=0.5, color="white", size=2.5) +
  ggtitle("Median IBU by State") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x="International Bitterness Unit (IBU)", y="States", fill="medianIBU") + 
  coord_flip() + 
  guides(fill=FALSE)
```

## Question 5A: Which state has the maximum alcoholic (ABV) beer?
```{r Question_5A, echo=TRUE}
# Getting the maximum ABV for each state
max_abv_by_state=bandb %>% group_by(bandb$State) %>% summarize(maxABV=max(ABV, na.rm=T))

# Assigning useful names to the 'median_abv_by_state' dataframe
names(max_abv_by_state)=c("State", "maxABV")

# Displaying partial results for the 'median_abv_by_state' dataframe
head(arrange(median_abv_by_state, desc(median_abv_by_state$medianABV)))
```

## Question 5B: Which state has the most bitter (IBU) beer?
```{r Question_5B, echo=TRUE}
# Getting the maximum IBU for each state
max_ibu_by_state=bandb %>% group_by(bandb$State) %>% summarize(maxibu=max(IBU, na.rm=T))

# Assigning useful names to the 'median_ibu_by_state' dataframe
names(max_ibu_by_state)=c("State", "maxIBU")

# Displaying partial results for the 'median_ibu_by_state' dataframe
head(arrange(median_ibu_by_state, desc(median_ibu_by_state$medianIBU)))
```

## Question 6: Summary statistics for the ABV variable.
```{r Question_6, echo=TRUE}
# Adding the 'purrr' library to the project
library(purrr)

# Generalized summary of the 'bandb$ABV' column
summary_abv=summary(bandb$ABV)

# Summarizing all of the 'bandb' dataframe by state
summary_by_state=bandb %>% split(bandb$State) %>% map(summary)

# Displaying partial results for the 'summary_by_state' list
head(summary_by_state, 1)
```

## Question 7A: Is there an apparent relationship between the bitterness of the beer and its alcoholic content?

|       <b>ANSWER:</b>
|
|               There does not appear to be a correlation between bitterness (IBU) and the alcoholic content (ABV).  
|               The scatter plot below shows the ABV / IBU of 1405 different beers in the United States.  If there was  
|               a correlation you would expect the data points to be closely centered around the linear regression line.
|               Instead the data is all over the place with no consistency.

## Question 7B: Draw a scatter plot.
```{r Question_7B, fig.width=10,fig.height=6, echo=TRUE}
# Creating a scatterplot of the ABV vs IBU data
ggplot(bandb, aes(x=bandb$ABV, y=bandb$IBU)) + geom_point(shape=10) + 
  geom_smooth(method=lm, se=FALSE) + 
  ggtitle("IBU vs ABV for Beers in the United States") + 
  theme(plot.title=element_text(hjust=0.5)) + 
  labs(x="Alcohol by Volume (ABV)", y="International Bitternss Unit (IBU)")
```


## Presentation Data:
```{r Presentation_Data, fig.width=10,fig.height=6, echo=TRUE}
mean_abv_by_state=bandb %>% group_by(bandb$State) %>% summarize(meanABV=mean(ABV, na.rm=T))
mean_ibu_by_state=bandb %>% group_by(bandb$State) %>% summarize(meanIBU=mean(IBU, na.rm=T))

names(mean_abv_by_state)=c("State", "meanABV")
names(mean_ibu_by_state)=c("State", "meanIBU")

write.csv(median_abv_by_state, file = "State_Median_ABV.csv")
write.csv(mean_abv_by_state, file = "State_Mean_ABV.csv")
write.csv(median_ibu_by_state, file = "State_Median_IBU.csv")
write.csv(mean_ibu_by_state, file = "State_Mean_IBU.csv")

```

|
|
|
|
|
|
|
|
| #######END OF DOCUMENT#######